Assigning Scouts to Optimal Patrols

### 2016-05-25 draft1.tex

%------------------------------------------------------------------------------
% End of journal.tex
%------------------------------------------------------------------------------

>>>>

%------------------------------------------------------------------------------
% End of journal.tex
%------------------------------------------------------------------------------
=======

\section*{Mathematical Model}
This problem visually lends itself well to the notion of a mathematical graph. Since there is an inherent sense of relationships between individuals (how much a scout likes another scout), it is appropriate to consider the scouts as vertices in a graph with directed edges representing how much the “head” scout likes the “tail” scout. Our goal, using this language, is to find a partition of the vertices in the graph that satisfies some sort of objective function given size constraints of the partitions. In our problem at hand, we consider a complete graph with 13 vertices and edge weights being either 0, 1, or -1, corresponding to indifference, approval, or disapproval respectively of the “head” scout with respect to the “tail” scout. The Scoutmaster specifically wants two groups out of these 13, and with the size constraints for a patrol considered, our goal becomes finding 6 vertices to be in one partition and the other 7 being in another. The total number of partitions becomes 13 choose 6, which is equivalent to 1,716. This is certainly within the scope of considering total enumeration of the partitions as a reasonable solution to the problem, therefore we will proceed accordingly and implement the calculation using Python and its associated libraries to represent the data in a format that is convenient to work with.

Now, a natural question for us to consider is what objective function would be most appropriate to optimize. A natural answer to this would be to sum the edge weights of all edges that connect vertices that are within partitions. This would make sense if separating enemies is just as important as joining friends, but this just might not be the case. In terms of weighing these factors, we defer to Scoutmaster Bruce’s experience with how past patrols fared with friends and foes in the mix. Therefore, we consider a family of possible objective functions, namely giving real numbered weights to instances of approving relationships and disapproving relationships, and select a few of these objective functions to solve for. We then present the collection of results from different objective functions to the Scoutmaster for him to see which one seems to most reasonably match his preferences.
\section*{Solution of Mathematical Problem}
We represent our directed graph as a 13 by 13 similarity matrix, or, as might be more fitting for the problem, an “affinity” matrix. Like below:%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\footnote{The scouts were enumerated like so: 1. Brandon, 2. Cameron, 3, Christian, 4. Colby, 5. Daniel, 6. Darwin, 7. Evan, 8. Jake, 9. Jordan, 10. Nathan, 11. Patrick, 12. Timmy, 13. Tommy.}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\includegraphics[scale=0.6]{solution1.JPG}

The NA value represents a hard constraint established by a scout’s parent (i.e., under no circumstances are these two to be in the same patrol). In our implementation we convert the NA value to $-\infty$. We designed an objective function that accepts two scouts who are assumed to be in different patrols and calculates the loss associated with such a placement. The loss is defined to be $L(x, y) = \text{max}(-2, x[y] + y[x])$ where $x[y]$ is $x$’s disposition towards $y$ and vise versa. For example, if $x$ dislikes $y$ and $y$ dislikes $x$, the loss is -2. If $x$ likes $y$ but $y$ dislikes $x$ the loss is 0. If $x$ and $y$ both like each other, the loss is 2. In mixed cases where the first scout is indifferent to another, but the other likes or dislikes the first scout, the loss is +1 or -1 respectively. We are trying to minimize the loss, so in applying our loss function over the entire dataset, our algorithm is equally averse to same-patrol mutual enmity as it is to dividing mutual friendship, half as averse to the indifference of one scout and the stronger inclination of another, and indifferent to polar opposite affinities within the same patrol. There is a question as to whether this is an accurate reflection of how individual scouts would view their own “loss” of any of the possible scenarios. As it turns out, our algorithm arrives at a very reasonable solution regardless. 
\section*{Results}
As mentioned earlier, the number of scouts we must partition is small enough to find an optimal solution by brute force. Our algorithm finds a partition of scouts into patrols of sizes 6 and 7 with a total loss of 0 as described in the table below:
\begin{table}[ht]
	\caption{}\label{eqtable}
	\renewcommand\arraystretch{1.5}
	\noindent\[
	\begin{array}{|c|c|}
	\hline
	\textbf{Patrol 1}&\textbf{Patrol 2}\\
	\hline
	\text{Brandon, Cameron, Colby,}&\text{Christian, Daniel, Jake,}\\
	\text{Darwin, Evan, Tommy}&\text{Jordan, Nathan, Patrick, Timmy}\\
	\hline
	\end{array}
	\]
\end{table}
\includegraphics[scale=0.65]{results1.JPG}
Of the undesirable outcomes of such a partition, we can spot a few:
\begin{enumerate}
	\item Tommy and Timmy both like each other and are in separate patrols. But if we move Timmy to Patrol 1 ($P1$), we break up the mutual friendship of Timmy and Nathan in $P2$. Suppose we moved both Timmy and Nathan to $P1$ and Brandon (who dislikes Nathan) to $P2$. But then Brandon loses two friends (Tommy and Cameron), gains one (Jordan), and now neither Nathan nor Timmy are in the same patrol as Daniel, whom they both like. Although, Brandon and Tommy find themselves in the awkward situation of having complete opposite sentiments towards each other. Perhaps this would be a worthwhile trade after all (and a larger, negative weight needs to be placed upon disfavor-favor relationships).

	\item As mentioned in the previous point, Tommy dislikes Brandon. Trading Tommy in $P1$ for Jordan in $P2$ seems promising, But we have overlooked the fact that Colby and Jordan are our NA pair, and cannot be in the same patrol under any circumstances. Trading Colby for Jordan is another option, but Colby will be missed by Darwin, Evan, Cameron and Tommy (In other words, the whole of P1 minus Brandon).
\end{enumerate}
On the whole, though, there are no obvious improvements that can be made to our algorithm’s partition. We find that this is an assignment that could have reasonably been arrived at by Scoutmaster Bruce (granted the group dynamics have not changed since the scouts were surveyed).  Furthermore, the solution took only 20 seconds to arrive at on a single core machine.

\section*{Improvements}
While our algorithm demonstrably works well on smaller groups of 13 people, we do not expect our brute force solution to be tractable on larger groups wherein greater than 2 partitions are required. As a possible scenario, consider a company of 150 people to be divided into fixed-size teams of 10 individuals each. There are $150! / (10!)^{15}$ possible partitions, a number proportional to $10^164$ - meaning we could instead use that computational time to enumerate the number of atoms in the universe… a tredecillion $(10^{78})$ times. Knowing this, perhaps it won’t surprise the reader that this particular graph partitioning problem in general graphs is NP-complete - though there do exist approximation algorithms.%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\footnote{B. W. Kernighan, S. Lin, \textit{An Efficient Heuristic Procedure for Partitioning Graphs}. Bell System Technical Journal. \textbf{49} (1970), 291--307. doi: 10.1002/j.1538-7305.1970.tb01770.x}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As scout enrollment increases, it becomes necessary to consider scenarios where we have both a large number of scouts in need of assignment and multiple patrols to choose from. Our algorithm could be extended to use brute force to find a globally optimal solution when it is deemed computationally feasible and to use an approximating algorithm to find a locally optimal solution otherwise.

%------------------------------------------------------------------------------
% End of journal.tex
%------------------------------------------------------------------------------
>>>>>>> 9771af696fbc57423120850744607162fcf171d5
